pseudobulk_CMP

### 2021-09-13 CMP/CMPSubset.Rmd
Creating new generate seurat object script and updating some other things

```{r setup}
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(clustree)
```

@@

store session info
```{r}
sessionInfo.filename <- paste0(projectName, "_sessionInfo.txt")
sink(sessionInfo.filename)
sessionInfo()
sink()
```

```{r}
source("../RFunctions/read_10XGenomics_data.R")
source("../RFunctions/PercentVariance.R")
source("../RFunctions/Mouse2Human_idconversion.R")
source ("../RFunctions/ColorPalette.R")
```

```{r warning=FALSE}
setwd("../../cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
data_file.list <- read_10XGenomics_data(sample.list = "CMPm2")
object.data <-Read10X(data_file.list)
```

@@

```{r}
plot1 <- FeatureScatter(cmp.object, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "orig.ident", pt.size = 0.01)
plot2 <- FeatureScatter(cmp.object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "orig.ident", pt.size = 0.01)
plot1 + plot2
```
## Filter data
remove low quality cells
require: nFeature_RNA between 200 and 4000 (inclusive)
require: percent.mt <=5

@@

Find variable features
```{r fig.width = 5, fig.height = 2}
cmp.object <- FindVariableFeatures(cmp.object, selection.method = "vst", nfeatures = 2000)
top10 <- head(VariableFeatures(cmp.object), 10)
plot1 <- VariableFeaturePlot(cmp.object)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2
```

## Scale data (linear transformation)
We don't have to worry about comparing library depths, so we'll just do normalization/Scale data

@@

```{r}
clust3.cells <- read.table(file = "../Seuratv1_clusterCellIDs/cluster3cellIDs.txt", col.names = "clust03")
clust3.cells <- sapply(clust3.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
clust17.cells <- read.table(file = "../Seuratv1_clusterCellIDs/cluster17cellIDs.txt", col.names = "clust17")
clust17.cells <- sapply(clust17.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
clust10.cells <- read.table(file = "../Seuratv1_clusterCellIDs/cluster10cellIDs.txt", col.names = "clust10")
clust10.cells <- sapply(clust10.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
clust11.cells <- read.table(file = "../Seuratv1_clusterCellIDs/cluster11cellIDs.txt", col.names = "clust11")
clust11.cells <- sapply(clust11.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
```

@@

@@

```
Plot UMAP

```{r}
for(x in c(0.5, 1, 1.5, 2, 2.5)){
	cmp.object <- FindClusters(cmp.object, resolution = x)
}
saveRDS(cmp.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
```

@@

### Clustree
what's the max resolution we can achieve while keepign clusters stable?
```{r fig.width = 10, fig.height = 5}
plot.title <- paste0(projectName, "_clustree_ndim", max(cmp.object@commands$RunUMAP.RNA.pca$dims))
my.clustree <- clustree(cmp.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability", exprs = "scale.data") + 
	scale_color_continuous(low = 'red3', high = 'white') + 
	ggtitle(plot.title)
png(filename = paste0(plot.title, ".png"), height = 800, width = 1600)
plot(my.clustree)
dev.off()

```
I think I'm liking res.1.0 from this. Although how much does this change if I use fewer PCs...

@@

# Total var 85%
## Neighborhood and umap
set total.var <- 85%
```{r}
tot.var <- percent.variance(cmp.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
ndims <- length(which(cumsum(tot.var) <= 85))

@@

```
Plot UMAP

```{r}
for(x in c(0.5, 1, 1.5, 2, 2.5)){
	cmp.object <- FindClusters(cmp.object, resolution = x)
saveRDS(cmp.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
}
```

### Clustree
what's the max resolution we can achieve while keepign clusters stable?
```{r fig.width = 10, fig.height = 5}
plot.title <- paste0(projectName, "_clustree_ndim", max(cmp.object@commands$RunUMAP.RNA.pca$dims))
my.clustree <- clustree(cmp.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability", exprs = "scale.data") + 
	scale_color_continuous(low = 'red3', high = 'white') + 
	ggtitle(plot.title)
png(filename = paste0(plot.title, ".png"), height = 800, width = 1600)
plot(my.clustree)
dev.off()
```

## Identify variable genes for new biomark
```{r}
length(cmp.object@assays$RNA@var.features)
```

### Gene profiles of clusters
set ident at res = 1 and get markers
```{r}
Idents(cmp.object) <- "RNA_snn_res.1"
cmp.allmarkers.res1 <- FindAllMarkers(cmp.object)
cmp.top100markers.res1 <- cmp.allmarkers.res1 %>% group_by(cluster) %>% top_n(n = 100, wt = abs(avg_log2FC))
```

```{r}
cmp.top100markers.res1 <- cmp.allmarkers.res1 %>% group_by(cluster) %>% top_n(n = 100, wt = abs(avg_log2FC))
cmp.top100markers.res1 <- cmp.top100markers.res1[cmp.top100markers.res1$p_val_adj <= 0.05, ]
for(cluster in unique(cmp.top100markers.res1$cluster)){
	num.hits <- nrow(cmp.top100markers.res1[cmp.top100markers.res1$cluster == cluster, ])
	print(paste("cluster", cluster, "has", num.hits, "genes"))
}
```

Find top.var	genes in cmp.top100markers.res1
```{r}
top.var <- cmp.object@assays$RNA@var.features[1:150]
# top.var
n.vargenes <- c()
for(gene in top.var){
	if(gene %in% cmp.top100markers.res1$gene){
		n.vargenes <- c(n.vargenes, gene)
	}
}
print(length(n.vargenes))
top.var[!(top.var %in% n.vargenes)]
```

Lets try to export these, with genes on rows and clusters on columns, and xpression value in data table
```{r}
var.df <- cmp.top100markers.res1[cmp.top100markers.res1$gene %in% n.vargenes,]
var.df <- subset.data.frame(var.df, select = c("gene", "cluster", "avg_log2FC"))
var.df <- reshape2::dcast(var.df, gene~cluster, value.var = "avg_log2FC")
xlsx::write.xlsx(var.df, file = paste0(projectName, "_dim", ndims, "_candididateBiomarkGenes.xlsx"), sheetName = "res1", append = TRUE, row.names = FALSE)

repeat for top.var	genes in cmp.top100markers.res1.5
```{r}
top.var <- cmp.object@assays$RNA@var.features[1:150]
Idents(cmp.object) <- "RNA_snn_res.1.5"
cmp.allmarkers.res1.5 <- FindAllMarkers(cmp.object)
cmp.top100markers.res1.5 <- cmp.allmarkers.res1.5 %>% group_by(cluster) %>% top_n(n = 100, wt = abs(avg_log2FC))
n.vargenes <- c()
for(gene in top.var){
	if(gene %in% cmp.top100markers.res1.5$gene){
		n.vargenes <- c(n.vargenes, gene)
	}
}
print(length(n.vargenes))
top.var[!(top.var %in% n.vargenes)]
```

Lets try to export these, with genes on rows and clusters on columns, and xpression value in data table
```{r}
var.df <- cmp.top100markers.res1.5[cmp.top100markers.res1.5$gene %in% n.vargenes,]
var.df <- subset.data.frame(var.df, select = c("gene", "cluster", "avg_log2FC"))
var.df <- reshape2::dcast(var.df, gene~cluster)
xlsx::write.xlsx(var.df, file = paste0(projectName, "_dim", ndims, "_candididateBiomarkGenes.xlsx"), sheetName = "res1.5", append = TRUE, row.names = FALSE)

```
For each resolution, what percentage of cells in each cluster are enriched for one of our clust.IDs?

# Total var 80%
## Neighborhood and umap
set total.var <- 80%
```{r}
tot.var <- percent.variance(cmp.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
ndims <- length(which(cumsum(tot.var) <= 80))

```
Plot UMAP

```{r}
for(x in c(0.5, 1, 1.5, 2, 2.5)){
	cmp.object <- FindClusters(cmp.object, resolution = x)
saveRDS(cmp.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
}
```

### Clustree
what's the max resolution we can achieve while keepign clusters stable?
```{r fig.width = 10, fig.height = 5}
plot.title <- paste0(projectName, "_clustree_ndim", max(cmp.object@commands$RunUMAP.RNA.pca$dims))
my.clustree <- clustree(cmp.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability", exprs = "scale.data") + 
	scale_color_continuous(low = 'red3', high = 'white') + 
	ggtitle(plot.title)
png(filename = paste0(plot.title, ".png"), height = 800, width = 1600)
plot(my.clustree)
dev.off()
```

# Total var 95%

## Neighborhood and umap
set total.var <- 95%
```{r}
tot.var <- percent.variance(cmp.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
ndims <- length(which(cumsum(tot.var) <= 95))

```
Plot UMAP

```{r}
for(x in c(0.5, 1, 1.5, 2, 2.5)){
	cmp.object <- FindClusters(cmp.object, resolution = x)
}
saveRDS(cmp.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
```

```{r}
cmp.object <- readRDS("CMP_dim41.RDS")
```

```{r fig.width = 10, fig.height = 5}
plot.title <- paste0(projectName, "_clustree_ndim", max(cmp.object@commands$RunUMAP.RNA.pca$dims))
my.clustree <- clustree(cmp.object, prefix = "RNA_snn_res.", node_colour = "sc3_stability", exprs = "scale.data") + 
	scale_color_continuous(low = 'red3', high = 'white') + 
	ggtitle(plot.title)
png(filename = paste0(plot.title, ".png"), height = 800, width = 1600)
plot(my.clustree)
dev.off()
```

@@

Create `FindAllMarkers()` lists for GSEA
```{r }
Idents(cmp.object) <- "RNA_snn_res.2.5"
res.2.5.allmarkers <- FindAllMarkers(cmp.object)
```

## Map HGNC symbols
```{r}
Mouse2HumanTable <- Mouse2Human(res.2.5.allmarkers$gene)

@@

calculate `FindMarkers()` that distinguish each cluster (might overlab between clusters)
```{r}
ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1", "RNA_snn_res.1.5", "RNA_snn_res.2", "RNA_snn_res.2.5", "clust.ID")
for(tested.ident in ident.list){
	object.copy <- cmp.object
	Idents(object.copy) <- tested.ident
	print(paste("testing", tested.ident))
	for (cluster in sort(as.numeric(levels(object.copy@meta.data[[tested.ident]])))){
		print(paste("looking at cluster", cluster))
		cluster.markers <- FindMarkers(object.copy, ident.1 = cluster)
		try(
			xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
											 file = paste0(projectName, "_FindMarkers_", gsub("RNA_snn_", "", tested.ident), ".xlsx"), 
											 sheetName = paste0("clst", cluster), 
											 col.names = TRUE, 
											 row.names = TRUE, 
											 append = TRUE)
		)	
	}
	remove(object.copy)
}
```

@@

# Compare @g hierarchcial clusteirng

>>>>

```{r setup}
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/CMP/")
# library(clustree)
```

@@

store session info
```{r }
# sink("CMP-v1.20210616.txt")
sessionInfo()
# sink()
```

```{r}
source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/read_10XGenomics_data.R")
source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/PercentVariance.R")
source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/Mouse2Human_idconversion.R")
```

```{r warning=FALSE}
setwd("~/Desktop/10XGenomicsData/cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
data_file.list <- read_10XGenomics_data(sample.list = "CMPm2")
object.data <-Read10X(data_file.list)
```

@@

```{r}
plot1 <- FeatureScatter(cmp.object, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "orig.ident", pt.size = 0.01)
plot2 <- FeatureScatter(cmp.object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "orig.ident", pt.size = 0.01)
plot1 + plot2
```

We don't have to worry about comparing library depths, so we'll just do normalization/Scale data

remove low quality cells
require: nFeature_RNA between 200 and 4000 (inclusive)
require: percent.mt <=5

@@

Find variable features
```{r fig.width = 5, fig.height = 2}
cmp.object <- FindVariableFeatures(cmp.object, selection.method = "vst", nfeatures = 2000)
top10 <- head(VariableFeatures(cmp.object), 10)
plot1 <- VariableFeaturePlot(cmp.object)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2

Scale data (linear transformation)

@@

```{r}
clust3.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster3cellIDs.txt", col.names = "clust03")
clust3.cells <- sapply(clust3.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
clust17.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster17cellIDs.txt", col.names = "clust17")
clust17.cells <- sapply(clust17.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
clust10.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster10cellIDs.txt", col.names = "clust10")
clust10.cells <- sapply(clust10.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
clust11.cells <- read.table(file = "Seuratv1_clusterCellIDs/cluster11cellIDs.txt", col.names = "clust11")
clust11.cells <- sapply(clust11.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
```

@@

### Color palette
```{r}
color.palette <- c(
	"coral",
	"chartreuse4",
	"goldenrod1",
	"cadetblue1",
	"burlywood",
	"brown",
	"brown1",
	"blue",
	"blue4",
	"azure3",
	"aquamarine",
	"antiquewhite",
	"cadetblue",
	"gold3",
	"black",
	"darkgreen",
	"deeppink",
	"darkviolet",
	"darkturquoise",
	"darkslategray",
	"darksalmon",
	"darkorchid1",
	"darkolivegreen2",
	"forestgreen",
	"dodgerblue",
	"green",
	"lightpink",
	"lightcoral",
	"khaki1",
	"maroon",
	"peru",
	"lightseagreen",
	"lightsalmon",
	"plum",
	"moccasin",
	"tan",
	"tan1", 
	"red", 
	"purple",
	"khaki4",
	"black", 
	"plum4"
)
```

@@

saveRDS(cmp.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
```
Plot UMAP

@@

@@

# Total var 95%
## Neighborhood and umap
set total.var <- 95%
```{r}
tot.var <- percent.variance(cmp.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
ndims <- length(which(cumsum(tot.var) <= 95))

@@

saveRDS(cmp.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
```
Plot UMAP

@@

Create `FindAllMarkers()` lists for GSEA
```{r}
Idents(cmp.object) <- "RNA_snn_res.2.5"
res.2.5.allmarkers <- FindAllMarkers(cmp.object)
```

Map HGNC symbols
```{r}
Mouse2HumanTable <- Mouse2Human(res.2.5.allmarkers$gene)

@@

calculate `FindMarkers()` that distinguish each cluster (might overlab between clusters)
```{r}
ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1", "RNA_snn_res.1.5", "RNA_snn_res.2", "RNA_snn_res.2.5", "clust.ID")
for(tested.ident in ident.list){
	for (cluster in sort(as.numeric(levels(cmp.object@meta.data[[tested.ident]])))){
	cluster.markers <- FindMarkers(cmp.object, ident.1 = cluster)
	xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
									 file = paste0(projectName, "_FindMarkers_", gsub("RNA_snn_", "", tested.ident), ".xlsx"), 
									 sheetName = paste0("clst", cluster), 
									 col.names = TRUE, 
									 row.names = TRUE, 
									 append = TRUE)
}
}
```

@@

# Cell cycle analysis
Just so we know what we're working with

```{r}
s.genes <- cc.genes.updated.2019$s.genes
g2m.genes <- cc.genes.updated.2019$g2m.genes
```

# Compare @ hierarchcial clusteirng

### 2021-09-13 pseudobulk_vs_bulk/pseudobulk_allPops.Rmd
Creating new generate seurat object script and updating some other things

## Update
### 2021.09.13
Finally got a method working for CMP vs. TotalSeq CMP (see pseudobulk_CMP.RMD). Will now clean up notebook and comparisons for both TotalSeq and ScriptSeq, and for all LSK, CMP, MEP, and GMP populations.

@@

Creating new pipeline using seurat v4.0.2 available 2021.06.08

Load libraries required for Seuratv4

## Load libraries
```{r setup}
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(SingleCellExperiment)
library(Matrix.utils)
library(DESeq2)
library(edgeR)
library(csaw)
library(tximport)
```

## Set global variables
```{r}
projectName <- "pseudobulk_allPops"
comparison_pop <- "CMP"
```

@@

## Load local scripts
```{r}
# source("../RFunctions/read_10XGenomics_data.R")
# source("../RFunctions/PercentVariance.R")
source("../RFunctions/Mouse2Human_idconversion.R")
source ("../RFunctions/ColorPalette.R")
```

# Load scRNASeq object

## Dim 25
Using CMP.rds object of dim=25, describing 85% of variation
```{r}
# CMP seurat object generated in "CMPSubset.Rmd"
seurat.object <- readRDS("CMP_dim25.RDS")
```
```{r fig.height=5, fig.width=15}
p <- FeaturePlot(seurat.object, features = c("Fli1", "Klf1"), blend = TRUE, pt.size = 1, combine = TRUE, keep.scale = "all")
plot(p)
png(filename = "CMP_Fli1Klf1Ratio.png", width = 2400, height = 400)
plot(p)
dev.off()
```

```{r fig.height=5, fig.width=15}
p <- FeaturePlot(seurat.object, features = c("Gata2", "Gata1"), blend = TRUE, pt.size = 1, combine = TRUE, keep.scale = "all")
plot(p)
png(filename = "CMP_Gata2Gata1Ratio.png", width = 2400, height = 400)
plot(p)
dev.off()
```

## Calculate pseudobulk gene expression

__ Based on tutorial @ https://hbctraining.github.io/scRNA-seq/lessons/pseudobulk_DESeq2_scrnaseq.html __
```{r}
# Extract raw counts and metadata to create SingleCellExperiment object

counts <- seurat.object@assays$RNA@counts

metadata <- seurat.object@meta.data

# Set up metadata as desired for aggregation and DE analysis
metadata$cluster_id <- factor(seurat.object@active.ident)

# Create single cell experiment object
sce <- SingleCellExperiment(assays = list(counts = counts), 
                           colData = metadata)

```

Investigate the object
```{r}
print(dim(counts(sce)))
print(dim(colData(sce)))
print(head(colData(sce)))
```

Note that QC was performed during Seurat analysis

```{r}
# sce <- perCellQCMetrics(sce)
```
## Aggregate counts
### Aggregate top-level counts

```{r}
groups <- colData(sce)[, "orig.ident"]
aggr.counts <- aggregate.Matrix(t(counts(sce)), groupings = groups, fun = "sum")

dim(aggr.counts)
```

```{r}
sc.dds <- DESeqDataSetFromMatrix(countData = aggr.counts, colData = metadata, design = ~orig.ident)

```

# Load bRNASeq data

```{r}
# import rsem.genes.results files

setwd("../../bRNASeq/")
file_list <- list.files(pattern = "^CMP", recursive = TRUE, include.dirs = TRUE)
names(file_list) <- lapply(file_list, function(x) gsub('.genes.results', '', x))
file_list
```

```{r}
mstxi <- tximport(files = paste0("../../bRNASeq/", file_list), type = 'rsem', txIn = FALSE, txOut = FALSE, geneIdCol = 3, )
colnames(mstxi$counts) <- gsub("^.*/", "", names(file_list))
```

```{r}
rna.counts <- merge(mstxi$counts, t(aggr.counts), by = "row.names", all = TRUE)
rna.counts[is.na(rna.counts)]<-0
rownames(rna.counts) <- rna.counts$Row.names
rna.counts <- rna.counts[, 2:ncol(rna.counts)]
head(rna.counts)
```

```{r}
dgelist_groups <- factor(c(unlist(stringr::str_extract_all(gsub("/.*", "", names(file_list)), pattern = "[a-zA-Z]+")), "SingleCell"))
levels(dgelist_groups)
```

```{r}
dgelist <- DGEList(counts = rna.counts, group = dgelist_groups)
dgelist
apply(dgelist$counts, 2, sum)
keep <- rowSums(cpm(dgelist) > 10) >=2
dgelist <- dgelist[keep,]
```

### Calculate library size
```{r}
dgelist$samples$lib.size <- colSums(dgelist$counts)
dgelist <- calcNormFactors(dgelist)
```

### Calculate dispersion

designMat
```{r}
designMat <- model.matrix(~ 0 + dgelist$samples$group)
colnames(designMat) <- levels(dgelist$samples$group)
designMat
```

estimate GLM dispersion and apply
```{r}
dgelist <- estimateGLMCommonDisp(dgelist, designMat)
dgelist <- estimateGLMTrendedDisp(dgelist, designMat, method = 'bin.spline')
dgelist <- estimateGLMTagwiseDisp(dgelist, designMat)
```

```{r}
fit <- glmQLFit(dgelist, designMat, robust = TRUE)
plotQLDisp(fit)
```

```{r}
qlf <- glmQLFTest(fit, coef = )
summary(decideTests(qlf))
```

#### rld on  individual samples

```{r}
rld <- rlog(round(dgelist$counts), blind = TRUE)
rld <- as.data.frame(rld)
rld.spear <- cor(rld, method = "spearman")
rld.melt <- reshape2::melt(rld)
colnames(rld.melt) <- c("sample", "rld")
```

```{r}
rld.spear
```

```{r}
ggplot(data = rld.melt, aes(x = sample, y = rld, fill = sample)) + geom_violin()
```

```{r}
ggplot(data = rld, aes(x = CMPm2, y = CMP1010)) + geom_point()
```

#### rld on pooled samples

```{r}
pool.counts <- dgelist$counts
pool.counts <- as.data.frame(pool.counts)
pool.counts$ScriptSeq <- (pool.counts$CMP1009 + pool.counts$CMP1010)/2
pool.counts$TruSeq <- (pool.counts$CMP443 + pool.counts$CMP448)/2
pool.counts <- subset(pool.counts, select = c(ScriptSeq, TruSeq, CMPm2))
pool.counts <- as.matrix(pool.counts)
```

```{r}
pool.rld <- rlog(round((pool.counts)), blind = TRUE)
pool.rld.spear <- cor(pool.rld, method = "spearman")
pool.rld.melt <- reshape2::melt(pool.rld)
colnames(pool.rld.melt) <- c("sample", "rld")
```

```{r}
pool.rld.spear
```

```{r}
ggplot(data = pool.rld.melt, aes(x = sample, y = rld, fill = sample)) + geom_violin()
```

```{r}
ggplot(data = pool.rld.melt, aes(x = CMPm2, y = ScriptSeq)) + geom_point()
```

Combine replicates from CPM table

```{r}
rna.cpm <- as.data.frame(cpm(dgelist, log = TRUE, normalized.lib.sizes = TRUE))
rna.cpm$ScriptSeq <- (rna.cpm$CMP1009 + rna.cpm$CMP1010)/2
rna.cpm$TruSeq <- (rna.cpm$CMP443 + rna.cpm$CMP448)/2
rna.cpm <- subset(rna.cpm, select = c(ScriptSeq, TruSeq, CMPm2))
```

plot correlation and scatter grams
```{r}
cpm.spear <- cor(rna.cpm, method = "spearman")
cpm.melt <- reshape2::melt(rld)
colnames(cpm.melt) <- c("sample", "rld")
```

```{r}
cpm.spear
```

## Create DESeq object from single cell data
```{r}
sc.dds <- DESeqDataSetFromMatrix(countData = aggr.counts, colData = metadata, design = ~orig.ident)
```

# Single cell/Bulk RNA-Seq concordance tutorial
Following a new (tutorial) [https://jdblischak.github.io/singlecell-qtl/sc-vs-bulk.html].

Calculate cpm from single cell umi matrix
```{r}
sc.mtx <- data.frame(rowSums(seurat.object@assays$RNA@counts))
sc.mtx <- sc.mtx/colSums(sc.mtx)*1000000
```
Tutorial states that only quantity directly cmoprable between bulk and scRNA-Seq is relative abundance (cites Pachter 2001).

read cmp bulk TPM
```{r}
# generate table with individual TPMs
bulk.path <- "../../bRNASeq/TotalSeq/"
file_list <- list.files(path = bulk.path, pattern = comparison_pop)
bulk.pop <- read.table(file = paste0(bulk.path, file_list[1]), header = TRUE)[1]
for(file in file_list){
	print(file)
	bulk.pop <- merge(bulk.pop, read.table(file = paste0(bulk.path, file), header = TRUE, sep = "\t")[, c(1,5)], by = "gene_id")
}

#calculate average tpm
bulk.pop[comparison_pop] <- rowMeans(x = bulk.pop[, c(2, 3)])

# clean and check df
bulk.pop <- bulk.pop[, -c(2, 3)]
head(bulk.pop)
```
```{r}
#calc log_rho's
bulk.rho <- log((bulk.pop[comparison_pop] + 1), 2) - log(colSums((bulk.pop[comparison_pop])+ 1), 2)
rownames(bulk.rho) <- bulk.pop$gene_id
colnames(bulk.rho) <- "bulk"
sc.rho <- log((sc.mtx + 1), 2) - log(colSums((sc.mtx+1)), 2)
colnames(sc.rho) <- "sc"
```

check that all values are numeric
```{r}
print(is.numeric(bulk.rho[,1]))
print(is.numeric(sc.rho[,1]))
```
## create concordance df
### set NA = 0
```{r}
concordance.NAasZero <- merge(bulk.rho, sc.rho, by = "row.names", all = TRUE)
concordance.NAasZero[is.na(concordance.NAasZero)] <- 0
rownames(concordance.NAasZero) <- concordance.NAasZero$Row.names
concordance.NAasZero <- concordance.NAasZero[, c(2, 3)]
```
#### plot data
```{r}
ggplot(data = concordance.NAasZero, aes(x = bulk, y = sc)) + 
	geom_point()  +
	ggtitle(paste0("concordance in ", comparison_pop))
```
```{r}
cor(concordance.NAasZero, method = "spearman")
```

### drop NA
```{r}
concordance.df <- merge(bulk.rho, sc.rho, by = "row.names", all = TRUE)
concordance.df <- na.omit(concordance.df)
rownames(concordance.df) <- concordance.df$Row.names
concordance.df <- concordance.df[, c(2, 3)]
```
```{r}
spearman.coeff <- cor(concordance.df, method = "spearman")[1, 2]
spearman.coeff
```
#### plot data
```{r}
ggplot(data = concordance.df, aes(x = bulk, y = sc)) + 
	geom_point()  +
	ggtitle(paste0("concordance in ", comparison_pop, " spearman coeff = ", round(spearman.coeff, digits = 3)))
```

>>>>

## Update
### 2021.09.13
Finally got a method working for CMP vs. TotalSeq CMP (see pseudobulk_CMP.RMD). Will now clean up notebook and comparisons for both TotalSeq and ScriptSeq, and for all LSK, CMP, MEP, and GMP populations.  
Following a new [tutorial][https://jdblischak.github.io/singlecell-qtl/sc-vs-bulk.html].

@@

## Load libraries
```{r setup}
library(dplyr)
library(Seurat)
library(ggplot2)
knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/pseudobulk_vs_bulk/")
```

## Set global variables
```{r}
projectName <- "pseudobulk_allPops"
comparison.pop <- "CMP"
export.graphs <- TRUE
```

@@

## Load local scripts
```{r}
# source("../RFunctions/read_10XGenomics_data.R")
# source("../RFunctions/PercentVariance.R")
# source("../RFunctions/Mouse2Human_idconversion.R")
source ("../RFunctions/ColorPalette.R")
```

# Load scRNASeq object
##Load raw object
Has feature and mt filters applied
```{r}
seurat.object <- readRDS(paste0("~/Desktop/10XGenomicsData/msAggr_scRNASeq/raw_seuratObjects/", comparison.pop, "_raw.rds"))
```

# Calculate cpm 
from single cell umi matrix
```{r}
sc.mtx <- data.frame(rowSums(seurat.object@assays$RNA@counts))
sc.mtx <- sc.mtx/colSums(sc.mtx)*1000000
```
_Tutorial states that only quantity directly cmoprable between bulk and scRNA-Seq is relative abundance (cites Pachter 2001)._

# Load bRNASeq data
## read cmp bulk TPM
```{r}
bulk.subfolders <- c("../../bRNASeq/TotalSeq/", "../../bRNASeq/ScriptSeq/")
bulk.path <- bulk.subfolders[1]
file_list <- list.files(path = bulk.path, pattern = comparison.pop)
bulk.pop <- read.table(file = paste0(bulk.path, file_list[1]), header = TRUE)[1]

for(file in file_list){
	print(file)
	bulk.pop <- merge(bulk.pop, read.table(file = paste0(bulk.path, file), header = TRUE, sep = "\t")[, c(1,5)], by = "gene_id")
}

#calculate average tpm
bulk.pop[comparison.pop] <- rowMeans(x = bulk.pop[, c(2, 3)])

# clean and check df
bulk.pop <- bulk.pop[, -c(2, 3)]
head(bulk.pop)
```
```{r}
#calc log_rho's
bulk.rho <- log((bulk.pop[comparison.pop] + 1), 2) - log(colSums((bulk.pop[comparison.pop])+ 1), 2)
rownames(bulk.rho) <- bulk.pop$gene_id
colnames(bulk.rho) <- "bulk"
sc.rho <- log((sc.mtx + 1), 2) - log(colSums((sc.mtx+1)), 2)
colnames(sc.rho) <- "sc"
```

check that all values are numeric
```{r}
print(is.numeric(bulk.rho[,1]))
print(is.numeric(sc.rho[,1]))
```
## create concordance df
```{r}
# drop NAs
concordance.df <- merge(bulk.rho, sc.rho, by = "row.names", all = TRUE)
concordance.df <- na.omit(concordance.df)
rownames(concordance.df) <- concordance.df$Row.names
concordance.df <- concordance.df[, c(2, 3)]
```
```{r}
spearman.coeff <- cor(concordance.df, method = "spearman")[1, 2]
spearman.coeff
```
## plot data
```{r}
cor.plot <- ggplot(data = concordance.df, aes(x = bulk, y = sc)) + 
	geom_point()  +
	ggtitle(paste0("concordance in ", comparison.pop, " spearman coeff = ", round(spearman.coeff, digits = 3)))
plot(cor.plot)
```

## print images
```{r}
if(export.graphs == TRUE){
	png(filename = paste0(comparison.pop, "_spearmanPlot.png"), height = 480, width = 800)
	plot(cor.plot)
	dev.off()
} else{
	print("Export graphs set to FALSE")
}
```

# Automate through bRNASeq data
```{r}
bulk.folders <- c("TotalSeq", "ScriptSeq")

for(sequencing.folder in bulk.folders){
	bulk.subfolders <- c(paste0("~/Desktop/10XGenomicsData/bRNASeq/", sequencing.folder, "/"))

	for(bulk.path in bulk.subfolders){
		bulk.path <- bulk.subfolders[1]
		file_list <- list.files(path = bulk.path, pattern = comparison.pop)
		bulk.pop <- read.table(file = paste0(bulk.path, file_list[1]), header = TRUE)[1]

		for(file in file_list){
			print(file)
			bulk.pop <- merge(bulk.pop, read.table(file = paste0(bulk.path, file), header = TRUE, sep = "\t")[, c(1,5)], by = "gene_id")
		}

		#calculate average tpm
		bulk.pop[comparison.pop] <- rowMeans(x = bulk.pop[, c(2, 3)])

		# clean and check df
		bulk.pop <- bulk.pop[, -c(2, 3)]
		head(bulk.pop)

		#calc log_rho's
		bulk.rho <- log((bulk.pop[comparison.pop] + 1), 2) - log(colSums((bulk.pop[comparison.pop])+ 1), 2)
		rownames(bulk.rho) <- bulk.pop$gene_id
		colnames(bulk.rho) <- "bulk"
		sc.rho <- log((sc.mtx + 1), 2) - log(colSums((sc.mtx+1)), 2)
		colnames(sc.rho) <- "sc"

		# check all entries are numeric
		print(is.numeric(bulk.rho[,1]))
		print(is.numeric(sc.rho[,1]))

		# drop NAs
		concordance.df <- merge(bulk.rho, sc.rho, by = "row.names", all = TRUE)
		concordance.df <- na.omit(concordance.df)
		rownames(concordance.df) <- concordance.df$Row.names
		concordance.df <- concordance.df[, c(2, 3)]

		# spearman plot
		spearman.coeff <- cor(concordance.df, method = "spearman")[1, 2]
		print(spearman.coeff)

		cor.plot <- ggplot(data = concordance.df, aes(x = bulk, y = sc)) + 
			geom_point()  +
			ggtitle(paste0("concordance in ", comparison.pop, " vs ", sequencing.folder, ": spearman coeff = ", round(spearman.coeff, digits = 3)))
		plot(cor.plot)

		if(export.graphs == TRUE){
			png(filename = paste0(comparison.pop, "_vs", sequencing.folder,"-spearmanPlot.png"), height = 480, width = 800)
			plot(cor.plot)
			dev.off()
		} else{
			print("Export graphs set to FALSE")
		}
	}

}
```
